# Multiple Regression Analysis: Inference

Also covered using [Python](http://solomonegash.com/econometrics/wooldridge_python/iexample4_py.html) and [Stata](http://www.solomonegash.com/woodridge1/iexample4.html)

## Example 4.1 Wage equation

```{r echo=TRUE, message=FALSE, warning=FALSE}
rm(list = ls())
library(wooldridge)
library(psych)
library(stargazer)
library(car)
options(width=120)
wage_mols <- lm(lwage ~ educ + exper + tenure + 1, data=wage1)
stargazer(wage_mols, type="text", align=TRUE)
```


## Example 4.2. Student performance
```{r echo=TRUE, message=FALSE, warning=FALSE}
math_lin_lin <- lm(math10 ~ totcomp + staff + enroll + 1, data=meap93)
stargazer(math_lin_lin, type="text", align=TRUE)
math_lin_log <- lm(math10 ~ ltotcomp + lstaff + lenroll + 1, data=meap93)
stargazer(math_lin_log, type="text", align=TRUE)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
stargazer(math_lin_lin, math_lin_log, type="text", align=TRUE)
```

## Example 4.3. Collage GPA
```{r}
gpa_mols <- lm(colGPA ~ hsGPA + ACT + skipped + 1, data=gpa1)
stargazer(gpa_mols, type="text", align=TRUE)
```

## Example 4.4. Campus crime & enrollment
```{r}
crime_ols <- lm(lcrime ~ lenroll + 1, data=campus)
stargazer(crime_ols, type="text", align=TRUE)
```


## Example 4.5. Housing prices
```{r}
ldist <- log(hprice2$dist)
hprice_mols <- lm(lprice ~ lnox +ldist + rooms + stratio + 1, data=hprice2)
stargazer(hprice_mols, type="text", align=TRUE)
```


## Example 4.6. Participation rates in 401k plans
```{r}
pension_mols <- lm(prate ~ mrate + age + totemp + 1, data=k401k)
stargazer(pension_mols, type="text", align=TRUE)
```


## Example 4.7. Job training 
(Only for the year 1987 and for nonunionized firms)
```{r}
jtrain87 <- subset(jtrain, jtrain$year==1987 & jtrain$union==0)
job_mols <- lm(lscrap ~ hrsemp + lsales + lemploy + 1, data=jtrain87)
stargazer(job_mols, type="text", align=TRUE)

```

## Example 4.8. RD and Sales

```{r}
lrd_mols <- lm(lrd ~ lsales + profmarg + 1, data=rdchem)
stargazer(lrd_mols, type="text", align=TRUE)
```

## Example 4.9. Parent's education on birth weight
```{r}
birth_mols <- lm(bwght ~ cigs + parity + faminc + motheduc + fatheduc + 1, data=bwght)
birth_ols <- lm(bwght ~ cigs + parity + faminc + 1, data=bwght)
stargazer(birth_mols, birth_ols, type="text", align=TRUE)
linearHypothesis(birth_mols, c("motheduc = 0", "fatheduc =0"))
```

## Exaploring further example 4.5.
```{r}
attend_ols_r <- lm(atndrte ~ priGPA + 1, data=attend)
attend_ols <- lm(atndrte ~ priGPA + ACT + 1, data=attend)
stargazer(attend_ols_r, attend_ols, type="text", align=TRUE)
```

## Example 4.10. Salary-pension tradeoff for teachers
```{r}
meap_ols1 <- lm(lsalary ~ bensal + 1, data=meap93)
meap_ols2 <- lm(lsalary ~ bensal + lenroll + lstaff + 1, data=meap93)
meap_ols3 <- lm(lsalary ~ bensal + lenroll + lstaff + droprate + gradrate + 1, data=meap93)
stargazer(meap_ols1, meap_ols2, meap_ols3,   type="text", align=TRUE)
```


